Bandit Multiclass Linear Classification: \\ Efficient Algorithms for the Separable Case

### 2019-01-23 ICML-submission/empirical-comparison.tex
Fixed experiments section

In this section, we provide an empirical evaluation on our algorithms, verifying their
effectiveness on linearly separable datasets.
We generate strongly and weakly linearly separable datasets with $K=3$ classes in
$\R^3$, where the third dimension of every example is always equal to $1$
(i.e. a \emph{constant feature}).
See Figures~\ref{figure:strongly-separable-dataset} and~\ref{figure:weakly-separable-dataset}
 for visualizations
of the two datasets, along with detailed descriptions of the two data distributions.

%~\autoref{figure:strongly-and-weakly-separable-datasets}
%For both datasets, the sample size is $T=5\times 10^6$ and the
%margin is $\gamma=0.05$.
%Projections of the examples from the two datasets onto their first two coordinates
% are shown in
%We added a third dimension
%to every example that is equal to $1$. In other words, we added the so-called
%\emph{constant feature}. The constant feature is necessary so that
%\autoref{figure:strongly-separable-dataset} is actually strongly linearly
%separable.

\begin{figure}[h]
\centering
\begin{subfigure}[b]{0.23\textwidth}
\captionsetup{justification=centering}
\begin{center}
\hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_points}
\caption{Strongly separable case}
\label{figure:strongly-separable-dataset}
\end{center}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.23\textwidth}
\captionsetup{justification=centering}
\centering
\hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_points}
\caption{Weakly separable case}
\label{figure:weakly-separable-dataset}
\end{subfigure}
\vspace*{-0.2cm}
\caption{Strongly and weakly separable datasets with $K=3$ classes. Here
 we show projections of the examples onto their first two coordinates, which
 lie in the unit ball of $\R^2$. Class 1 is depicted red. Classes 2 and
3 are depicted green and blue, respectively. $80\%$ of the examples belong to class 1,
$10\%$ belong to class 2 and $10\%$ belong to class 3. Class 1
occupies the angle interval $[-15^\circ, 15^\circ]$, while classes 2 and 3
occupy angle intervals $[15^\circ, 180^\circ]$ and $[-180^\circ, -15^\circ]$
respectively. The examples are strongly and weakly linearly separable with a margin of $\gamma=0.05$, respectively.}
\label{figure:strongly-and-weakly-separable-datasets}
\end{figure}

%We removed points that lie within margin $\gamma=0.05$ of the
%linear separator.

We implement
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples},
Algorithm~\ref{algorithm:kernelized} with
rational kernel~\eqref{equation:rational-kernel}, and the \textsc{Banditron} algorithm.
We evaluate these algorithms on the above two datasets, with a time horizon of
$T=5\times 10^6$.
\textsc{Banditron}
has exploration rate parameter $\epsilon$, for which we tried values $\epsilon
\in \{0.02, 0.01, 0.005, 0.002, 0.001, 0.0005 \}$.
Since all three
algorithms are randomized, we run each algorithm $20$ times.
The average cumulative number
of mistakes up to round $t$ as a function of $t$ are shown in
Figures~\ref{figure:number-of-mistakes-strongly-separable-dataset}
and~\ref{figure:number-of-mistakes-weakly-separable-dataset}.

We can see that there is a tradeoff in the setting of the exploration rate $\epsilon$
for \textsc{Banditron}. For a large $\epsilon$, \textsc{Banditron} suffers from the cost of exploration, whereas for small $\epsilon$, its model cannot be updated quickly enough.
As expected, Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} has a small number of mistakes in the strongly linearly separable setting, while having a large number
of mistakes in the weakly linearly separable setting, due to the limited representation power of linear classifiers.
In contrast,
Algorithm~\ref{algorithm:kernelized} with rational kernel has a small number of mistakes in both settings, exhibiting strong adaptivity guarantees. Appendix~\ref{section:supp-to-experiment} shows the decision boundaries
that each of the algorithms learns.

%This problem is exacerbated when a class of large proportion occupies a small region
%(e.g. class 1).
%In contrast, our algorithms make
%significant progress on every mistake it makes, leading to a finite mistake
%bound.
%Striking a balance between the two extremes leads to the
%$\Theta(\sqrt{T})$ number of mistakes.

\begin{figure}
\centering
\includegraphics[width=0.42\textwidth]{figures/strong3}
\caption{The cumulative number of mistakes versus the number of rounds,
in the strongly separable distribution of \autoref{figure:strongly-separable-dataset} for various algorithms.}
\label{figure:number-of-mistakes-strongly-separable-dataset}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.42\textwidth]{figures/weak3}
\caption{The cumulative number of mistakes versus the number of rounds,
in the weakly separable distribution of \autoref{figure:weakly-separable-dataset} for various algorithms.}
\label{figure:number-of-mistakes-weakly-separable-dataset}
\end{figure}

>>>>

In this section, we provide an empirical evaluation on our algorithms, verifying
their effectiveness on linearly separable datasets. We generated strongly and
weakly linearly separable datasets with $K=3$ classes in $\R^3$ i.i.d. from two
data distributions. Figures~\ref{figure:strongly-separable-dataset}
and~\ref{figure:weakly-separable-dataset} show visualizations of the two
datasets, along with detailed descriptions of the distributions.

\begin{figure}[h]
\centering
\begin{subfigure}[b]{0.23\textwidth}
\captionsetup{justification=centering}
\begin{center}
\hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_points}
\caption{Strongly separable case}
\label{figure:strongly-separable-dataset}
\end{center}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.23\textwidth}
\captionsetup{justification=centering}
\centering
\hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_points}
\caption{Weakly separable case}
\label{figure:weakly-separable-dataset}
\end{subfigure}
\vspace*{-0.2cm}
\caption{Strongly and weakly linearly separable datasets in $\R^3$ with $K=3$
classes and $T=5\times 10^6$ examples. Here we show projections of the examples
onto their first two coordinates, which lie in the ball of radius $1/\sqrt{2}$
centered at the origin. The third coordinate is $1/\sqrt{2}$ for all examples.
Class 1 is depicted red. Classes 2 and 3 are depicted green and blue,
respectively. $80\%$ of the examples belong to class 1, $10\%$ belong to class 2
and $10\%$ belong to class 3. Class 1 lies in the angle interval $[-15^\circ,
15^\circ]$, while classes 2 and 3 lie in the angle intervals $[15^\circ,
180^\circ]$ and $[-180^\circ, -15^\circ]$ respectively. The examples are
strongly and weakly linearly separable with a margin of $\gamma=0.05$,
respectively. (Examples lying within margin $\gamma$ of the linear separators
were rejected during sampling.)}
\label{figure:strongly-and-weakly-separable-datasets}
\end{figure}

We implemented
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples},
Algorithm~\ref{algorithm:kernelized} with rational
kernel~\eqref{equation:rational-kernel}, and the \textsc{Banditron} algorithm.
We evaluated these algorithms on the two datasets. \textsc{Banditron} has an
exploration rate parameter, for which we tried values $0.02, 0.01, 0.005, 0.002,
0.001, 0.0005$. Since all three algorithms are randomized, we run each algorithm
$20$ times. The average cumulative number of mistakes up to round $t$ as a
function of $t$ are shown in
Figures~\ref{figure:number-of-mistakes-strongly-separable-dataset}
and~\ref{figure:number-of-mistakes-weakly-separable-dataset}.

We can see that there is a tradeoff in the setting of the exploration rate for
\textsc{Banditron}. With large exploration parameter, \textsc{Banditron} suffers
from over-exploration, whereas with small exploration parameter, its model
cannot be updated quickly enough. As expected,
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} has
a small number of mistakes in the strongly linearly separable setting, while
having a large number of mistakes in the weakly linearly separable setting, due
to the limited representation power of linear classifiers. In contrast,
Algorithm~\ref{algorithm:kernelized} with rational kernel has a small number of
mistakes in both settings, exhibiting strong adaptivity guarantees.
Appendix~\ref{section:supp-to-experiment} shows the decision boundaries that
each of the algorithms learns.

\begin{figure}
\centering
\includegraphics[width=0.42\textwidth]{figures/strong3}
\caption{Average cumulative number of mistakes of various algorithms versus the
number of rounds for the strongly linearly separable dataset of
\autoref{figure:strongly-separable-dataset}.}
\label{figure:number-of-mistakes-strongly-separable-dataset}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.42\textwidth]{figures/weak3}
\caption{Average cumulative number of mistakes of various algorithms versus the
number of rounds for the weakly linearly separable dataset of
\autoref{figure:weakly-separable-dataset}.}
\label{figure:number-of-mistakes-weakly-separable-dataset}
\end{figure}
